# Time series graphics {#ch-graphics}

The first thing to do in any data analysis task is to plot the data. Graphs enable many features of the data to be visualized, including patterns, unusual observations, changes over time, and relationships between variables. The features that are seen in plots of the data must then be incorporated, as far as possible, into the forecasting methods to be used. Just as the type of data determines what forecasting method to use, it also determines what graphs are appropriate.

But before we produce graphs, we need to set up our time series in R.

## `ts` objects

A time series can be thought of as a list of numbers, along with some information about what times those numbers were recorded. This information can be stored as a `ts` object in R.

Suppose you have annual observations for the last few years:

```{r tstable, echo=FALSE}
x <- c(123,39,78,52,110)
yr <- 2012:2016
knitr::kable(data.frame(Year=yr,Observation=x), booktabs=TRUE)
```

We turn this into a `ts` object using the `ts` function:

```r
y <- ts(c(123,39,78,52,110), start=2012)
```

If you have annual data, with one observation per year, you only need to provide the starting year (or the ending year).

For observations that are more frequent than once per year, you simply add a `frequency` argument. For example, if your monthly data is already stored as a numerical vector `z`, then it can be converted to a `ts` object like this:

```r
y <- ts(z, start=2003, frequency=12)
```

Almost all of the data used in this book is already stored as `ts` objects. But if you want to work with your own data, you will need to use the `ts` function before proceeding with the analysis.

### Frequency of a time series {-}

The "frequency" is the number of observations before the seasonal pattern repeats.^[This is the opposite of the definition of frequency in physics, or in Fourier analysis, where this would be called the "period".] When using the `ts()` function in R, the following choices should be used.

```{r freqtable, echo=FALSE}
tab <- data.frame(Data=c("Annual","Quarterly","Monthly","Weekly"),
  frequency=c(1,4,12,52))
knitr::kable(tab, booktabs=TRUE)
```

Actually, there are not 52 weeks in a year, but 365.25/7 = `r format(365.25/7,digits=4)` on average, allowing for a leap year every fourth year. But most functions which use `ts` objects require integer frequency.

If the frequency of observations is greater than once per week, then there is usually more than one way of handling the frequency. For example, data with daily observations might have a weekly seasonality (frequency=7) or an annual seasonality (frequency=365.25). Similarly, data that are observed every minute might have an hourly seasonality (frequency=60), a daily seasonality (frequency=24x60=1440), a weekly seasonality (frequency=24x60x7=10080) and an annual seasonality (frequency=24x60x365.25=525960). If you want to use a `ts` object, then you need to decide which of these is the most important.

In chapter \@ref(ch-advanced) we will look at handling these types of multiple seasonality, without having to choose just one of the frequencies.


##Time plots

For time series data, the obvious graph to start with is a time plot. That is, the observations are plotted against the time of observation, with consecutive observations joined by straight lines. Figure \@ref(fig:ansett) below shows the weekly economy passenger load on Ansett Airlines between Australia's two largest cities.

```{r ansett, fig.cap="Weekly economy passenger load on Ansett Airlines."}
autoplot(melsyd[,"Economy.Class"]) +
  ggtitle("Economy class passengers: Melbourne-Sydney") +
  xlab("Year") + ylab("Thousands")
```

We will use the `autoplot` command frequently. It automatically produces an appropriate plot of whatever you pass to it in the first argument. In this case, it recognizes `melsyd[,"Economy.Class"]` as a time series and produces a time plot.

The time plot immediately reveals some interesting features.

-   There was a period in 1989 when no passengers were carried --- this was due to an industrial dispute.
-   There was a period of reduced load in 1992. This was due to a trial in which some economy class seats were replaced by business class seats.
-   A large increase in passenger load occurred in the second half of 1991.
-   There are some large dips in load around the start of each year. These are due to holiday effects.
-   There is a long-term fluctuation in the level of the series which increases during 1987, decreases in 1989, and increases again through 1990 and 1991.
-   There are some periods of missing observations.

Any model will need to take all these features into account in order to effectively forecast the passenger load into the future.

A simpler time series is shown in Figure \@ref(fig:a10).

```{r a10, fig.cap="Monthly sales of antidiabetic drugs in Australia."}
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") + xlab("Year")
```

Here, there is a clear and increasing trend. There is also a strong seasonal pattern that increases in size as the level of the series increases. The sudden drop at the start of each year is caused by a government subsidisation scheme that makes it cost-effective for patients to stockpile drugs at the end of the calendar year.   
Any forecasts of this series would need to capture the seasonal pattern, and the fact that the trend is changing slowly.

##Time series patterns {#tspatterns}

In describing these time series, we have used words such as "trend" and "seasonal" which need to be defined more carefully.

Trend
  : A *trend* exists when there is a long-term increase or decrease in the data. It does not have to be linear. Sometimes we will refer to a trend as "changing direction", when it might go from an increasing trend to a decreasing trend. There is a trend in the antidiabetic drug sales data shown in Figure \@ref(fig:a10).

Seasonal
  : A *seasonal* pattern occurs when a time series is affected by seasonal factors such as the time of the year or the day of the week. Seasonality is always of a fixed and known frequency. The monthly sales of antidiabetic drugs above shows seasonality which is induced partly by the change in the cost of the drugs at the end of the calendar year.

Cyclic
  : A *cycle* occurs when the data exhibit rises and falls that are not of a fixed frequency. These fluctuations are usually due to economic conditions, and are often related to the "business cycle".  The duration of these fluctuations is usually at least 2 years.

Many people confuse cyclic behaviour with seasonal behaviour, but they are really quite different. If the fluctuations are not of a fixed frequency then they are cyclic; if the frequency is unchanging and associated with some aspect of the calendar, then the pattern is seasonal. In general, the average length of cycles is longer than the length of a seasonal pattern, and the magnitudes of cycles tend to be more variable than the magnitudes of seasonal patterns. Cycles and seasonality are discussed further in Section \@ref(sec-6-1-TSpatterns).

Many time series include trend, cycles and seasonality. When choosing a forecasting method, we will first need to identify the time series patterns in the data, and then choose a method that is able to capture the patterns properly.

The following four examples show different combinations of the above components.

```{r 6-decomp1, echo=FALSE, fig.cap="Four examples of time series showing different patterns."}
smallfonts <- theme(text = element_text(size = 9),
            axis.text = element_text(size=8))
p1 <- autoplot(hsales) + smallfonts +
        xlab("Year") + ylab("millions") +
        ggtitle("Sales of new one-family houses, USA")
p2 <- autoplot(ustreas) + smallfonts +
        xlab("Day") + ylab("Number") +
        ggtitle("US treasury bill contracts")
p3 <- autoplot(qauselec) + smallfonts +
        xlab("Year") + ylab("billion kWh") +
        ggtitle("Australian quarterly electricity production")
p4 <- autoplot(diff(dj)) + smallfonts +
        xlab("Day") + ylab("Change in index") +
        ggtitle("Dow Jones index")
gridExtra::grid.arrange(p1,p2,p3,p4,ncol=2)
```

  1. The monthly housing sales (top left) show strong seasonality within each year, as well as some strong cyclic behaviour with a period of about 6--10 years. There is no apparent trend in the data over this period.

  2. The US treasury bill contracts (top right) show results from the Chicago market for 100 consecutive trading days in 1981. Here there is no seasonality, but an obvious downward trend. Possibly, if we had a much longer series, we would see that this downward trend is actually part of a long cycle, but when viewed over only 100 days it appears to be a trend.

  3. The Australian monthly electricity production (bottom left) shows a strong increasing trend, with strong seasonality. There is no evidence of any cyclic behaviour here.

  4. The daily change in the Dow Jones index (bottom right) has no trend, seasonality or cyclic behaviour. There are random fluctuations which do not appear to be very predictable, and no strong patterns that would help with developing a forecasting model.

##Seasonal plots

A seasonal plot is similar to a time plot except that the data are plotted against the individual "seasons" in which the data were observed. An example is given below showing the antidiabetic drug sales.


```{r seasonplot1, fig.cap="Seasonal plot of monthly antidiabetic drug sales in Australia.", out.width="90%"}
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") + ggtitle("Seasonal plot: antidiabetic drug sales")
```

These are exactly the same data as were shown earlier, but now the data from each season are overlapped. A seasonal plot allows the underlying seasonal pattern to be seen more clearly, and is especially useful in identifying years in which the pattern changes.

In this case, it is clear that there is a large jump in sales in January each year. Actually, these are probably sales in late December as customers stockpile before the end of the calendar year, but the sales are not registered with the government until a week or two later. The graph also shows that there was an unusually small number of sales in March 2008 (most other years show an increase between February and March). The small number of sales in June 2008 is probably due to incomplete counting of sales at the time the data were collected.

A useful variation on the seasonal plot uses polar coordinates. Setting `polar=TRUE` makes the time series axis circular rather than horizontal, as shown below.

```{r seasonplot2, fig.cap="Polar seasonal plot of monthly antidiabetic drug sales in Australia.", out.width="90%"}
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") + ggtitle("Polar seasonal plot: antidiabetic drug sales")
```

##Seasonal subseries plots

An alternative plot that emphasises the seasonal patterns is where the data for each season are collected together in separate mini time plots.

```{r subseriesplot, fig.cap="Seasonal subseries plot of monthly antidiabetic drug sales in Australia."}
ggsubseriesplot(a10) + ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```

The horizontal lines indicate the means for each month. This form of plot enables the underlying seasonal pattern to be seen clearly, and also shows the changes in seasonality over time. It is especially useful in identifying changes within particular seasons. In this example, the plot is not particularly revealing; but in some cases, this is the most useful way of viewing seasonal changes over time.

##Scatterplots

The graphs discussed so far are useful for visualizing individual time series. It is also useful to explore relationships *between* time series.

Figure \@ref(fig:edemand) shows two time series: half-hourly electricity demand (in GigaWatts) and temperature (in degrees Celsius), for 2014 in Victoria, Australia. The temperatures are for Melbourne, the largest city in Victoria, while the demand values are for the entire state.

```{r edemand, fig.cap="Half hourly electricity demand and temperatures in Victoria, Australia, for 2014."}
month.breaks <- cumsum(c(0,31,28,31,30,31,30,31,31,30,31,30,31)*48)
autoplot(elecdemand[,c(1,3)], facet=TRUE) +
  xlab("Year: 2014") + ylab("") +
  ggtitle("Half-hourly electricity demand: Victoria, Australia") +
  scale_x_continuous(breaks=2014+month.breaks/max(month.breaks), 
    minor_breaks=NULL, labels=c(month.abb,month.abb[1]))
```

We can study the relationship between demand and temperature by plotting one series against the other.

```{r edemand2, fig.cap="Half-hourly electricity demand plotted against temperature for 2014 in Victoria, Australia."}
qplot(Temperature, Demand, data=as.data.frame(elecdemand)) +
  ylab("Demand (GW)") + xlab("Temperature (Celsius)")
```

This scatterplot helps us to visualize the relationship between the variables. It is clear that high demand occurs when temperatures are high due to the effect of air-conditioning. But there is also a heating effect, where demand increases for very low temperatures. 



### Scatterplot matrices {-}

When there are several potential predictor variables, it is useful to plot each variable against each other variable.  Consider the eight time series shown in Figure \@ref(fig:vntimeplots).

```{r vntimeplots, fig.cap="Quarterly visitor nights for various regions of Australia.", fig.asp=1}
autoplot(vn, facets=TRUE) +
  ylab("Number of visitor nights each quarter")
```

To see the relationships between these eight time series, we can plot each time series against the others. These plots can be arranged in a scatterplot matrix, as shown in Figure \@ref(fig:ScatterMatrixch2).

```{r ScatterMatrixch2, echo=TRUE, fig.cap="A scatterplot matrix of the quarterly visitor nights in eight regions of Australia.", fig.asp=1, fig.height=10, fig.width=10}
vn %>% as.data.frame() %>% GGally::ggpairs()
```

For each panel, the variable on the vertical axis is given by the variable name in that row, and the variable on the horizontal axis is given by the variable name in that column. There are many options available to produce different plots within each panel. In the default version, the correlations are shown in the upper right half of the plot, while the scatterplots are shown in the lower half. On the diagonal are shown density plots.

The value of the scatterplot matrix is that it enables a quick view of the relationships between all pairs of variables. Outliers can also be seen. In this example, there is one unusually high quarter for Sydney, corresponding to the 2000 Sydney Olympics.


##Lag plots

Figure \@ref(fig:beerlagplot) displays scatterplots of quarterly Australian beer production, where the horizontal axis shows lagged values of the time series. Each graph shows $y_{t}$ plotted against $y_{t-k}$ for different values of $k$.

```{r beerlagplot, fig.cap="Lagged scatterplots for quarterly beer production.", fig.asp=1}
beer2 <- window(ausbeer, start=1992)
gglagplot(beer2)
```

Here the colours indicate the quarter of the variable on the vertical axis.
The relationship is strongly positive at lags 4 and 8, reflecting the strong quarterly seasonality in the data.  The negative relationship seen for lags 2 and 6 occurs because peaks (in Q4) are plotted against troughs (in Q2)

The `window` function used here is very useful when extracting a portion of a time series. In this case, we have extracted the data from `ausbeer`, beginning in 1992.


##Autocorrelation

Just as correlation measures the extent of a linear relationship between two variables, autocorrelation measures the linear relationship between *lagged values* of a time series.

There are several autocorrelation coefficients, corresponding to each panel in the lag plot. For example, $r_{1}$ measures the relationship between $y_{t}$ and $y_{t-1}$, $r_{2}$ measures the relationship between $y_{t}$ and $y_{t-2}$, and so on.

The value of $r_{k}$ can be written as
$$
 r_{k} = \frac{\sum\limits_{t=k+1}^T (y_{t}-\bar{y})(y_{t-k}-\bar{y})}
 {\sum\limits_{t=1}^T (y_{t}-\bar{y})^2},
$$
where $T$ is the length of the time series.

The first nine autocorrelation coefficients for the beer production data are given in the following table.

```{r beeracftable, echo=FALSE, dependson="beerlagplot"}
beeracf <- matrix(acf(c(beer2), lag.max=9,
                      plot=FALSE)$acf[-1,,1], nrow=1)
colnames(beeracf) <- paste("$r_",1:9,"$",sep="")
knitr::kable(beeracf, booktabs=TRUE, format="pandoc",
             align="c", digits=3,
             format.args=list(nsmall=3))
```

These correspond to the nine scatterplots in Figure \@ref(fig:beerlagplot). The autocorrelation coefficients are normally plotted to form the *autocorrelation function* or ACF. The plot is also known as a *correlogram*.

```{r beeracf, fig.cap="Autocorrelation function of quarterly beer production.", fig.asp=0.35, dependson="beerlagplot"}
ggAcf(beer2)
```

In this graph:

-   $r_{4}$ is higher than for the other lags. This is due to the seasonal pattern in the data: the peaks tend to be four quarters apart and the troughs tend to be two quarters apart.
-   $r_{2}$ is more negative than for the other lags because troughs tend to be two quarters behind peaks.
-   The dashed blue lines indicate whether the correlations are significantly different from zero. These are explained in Section \@ref(wn).

### Trend and seasonality in ACF plots {-}


When data have a trend, the autocorrelations for small lags tend to be large and positive because observations nearby in time are also nearby in size. So the ACF of trended time series tend to have positive values that slowly decrease as the lags increase.

When data are seasonal, the autocorrelations will be larger for the seasonal lags (at multiples of the seasonal frequency) than for other lags.

When data are both trended and seasonal, you see a combination of these effects, as illustrated in Figure \@ref(fig:acfelec).


```{r aelec, echo=TRUE, fig.cap="Monthly Australian electricity demand from 1980--1995."}
aelec <- window(elec, start=1980)
autoplot(aelec) + xlab("Year") + ylab("GWh")
```

```{r acfelec, echo=TRUE, fig.cap="ACF of monthly Australian electricity demand.", fig.asp=0.35, dependson="aelec"}
ggAcf(aelec, lag=48)
```

The slow decrease in the ACF as the lags increase is due to the trend, while the "scalloped" shape is due the seasonality.

##White noise {#wn}

Time series that show no autocorrelation are called "white noise". Figure \@ref(fig:wnoise) gives an example of a white noise series.

```{r wnoise, fig.cap="A white noise time series."}
set.seed(30)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
```

```{r wnoiseacf, fig.cap="Autocorrelation function for the white noise series.", fig.asp=0.35, dependson="wnoise"}
ggAcf(y)
```

For white noise series, we expect each autocorrelation to be close to zero. Of course, they will not be exactly equal to zero as there is some random variation. For a white noise series, we expect 95% of the spikes in the ACF to lie within $\pm 2/\sqrt{T}$ where $T$ is the length of the time series. It is common to plot these bounds on a graph of the ACF (the blue dashed lines above). If one or more large spikes are outside these bounds, or if substantially more than 5% of spikes are outside these bounds, then the series is probably not white noise.

In this example, $T=50$ and so the bounds are at $\pm 2/\sqrt{50} = \pm `r format(2/sqrt(50),digits=2)`$. All of the autocorrelation coefficients lie within these limits, confirming that the data are white noise.


## Exercises {#ex-graphics}

1. Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.
  
    a. Use `autoplot` to plot each of these in separate plots.
    b. What is the frequency of each commodity series? Hint: apply the `frequency()` function.
    c. Use `which.max()` to spot the outlier in the `gold` series. Which observation was it?

2. Download some data from [OTexts.org/fpp2/extrafiles/tute1.csv](extrafiles/tute1.csv). Open the file `tute1.csv` in Excel (or some other spreadsheet application) and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

    a.  You can read the data into R with the following script:

        ```r
        tute1 <- read.csv("tute1.csv", header=TRUE)
        View(tute1)
        ```

    b.  Convert the data to time series

        ```r
        mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
        ```

        (The `[,-1]` removes the first column which contains the quarters as we don't need them now.)

    c.  Construct time series plots of each of the three series
    
        ```r
        autoplot(mytimeseries, facets=TRUE)
        ```

        Check what happens when you don't include `facets=TRUE`.

3. Download some monthly Australian retail data from [OTexts.org/fpp2/extrafiles/retail.xlsx](extrafiles/retail.xlsx). These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

    a. You can read the data into R with the following script:

        ```r
        retaildata <- readxl::read_excel("retail.xlsx", skip=1)
        ```

        You may need to first install the `readxl` package. The second argument (`skip=1`) is required because the Excel sheet has two header rows.

    b. Select one of the time series as follows (but replace the column name with your own chosen column):

        ```r
        myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
        ```

    c. Explore your chosen retail time series using the following functions:

        `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

        Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

4. Repeat for the following series:

    `bicoal`, `chicken`, `dole`, `usdeaths`, `bricksq`, `lynx`, `ibmclose`, `sunspotarea`, `hsales`, `hyndsight` and `gasoline`.

    Use the help files to find out what the series are.

5. The `arrivals` data set comprises quarterly international arrivals (in thousands) to Australia from Japan, New Zealand, UK and the US. Use `autoplot` and `ggseasonplot` to compare the differences between the arrivals from these four countries. Can you identify any unusual observations?

6. The following time plots and ACF plots correspond to four different time series. Your task is to match each  time plot in the first row with one of the ACF plots in the second row.

    ```{r acfguess, fig.asp=0.5, fig.width=12, echo=FALSE, message=FALSE, warning=FALSE}
    tp1 <- autoplot(cowtemp) + xlab("") + ylab("chirps per minute") +
      ggtitle("1. Daily temperature of cow")
    tp2 <- autoplot(USAccDeaths/1e3) + xlab("") + ylab("thousands") +
      ggtitle("2. Monthly accidental deaths")
    tp3 <- autoplot(AirPassengers) + xlab("") + ylab("thousands") +
      ggtitle("3. Monthly air passengers")
    tp4 <- autoplot(mink/1e3) + xlab("") + ylab("thousands") +
      ggtitle("4. Annual mink trappings")
    acfb <- ggAcf(cowtemp, ci=0) + xlab("") + ggtitle("B") + ylim(-0.4,1)
    acfa <- ggAcf(USAccDeaths, ci=0) + xlab("") + ggtitle("A") + ylim(-0.4,1)
    acfd <- ggAcf(AirPassengers, ci=0) + xlab("") + ggtitle("D") + ylim(-0.4,1)
    acfc <- ggAcf(mink, ci=0) + xlab("") + ggtitle("C") + ylim(-0.4,1)
    gridExtra::grid.arrange(tp1,tp2,tp3,tp4,
                            acfa,acfb,acfc,acfd,nrow=2)
    ```

7. The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots from Figures \@ref(fig:wnoise) and \@ref(fig:wnoiseacf).

## Further reading

 * @Cleveland1993 is a classic book on the principles of visualization for data analysis. While it is more than 20 years old, the ideas are timeless.
 * @Unwin2015 is a modern introduction to graphical data analysis using R. It does not have much information on time series graphics, but plenty of excellent general advice on using graphics for data analysis.

