## Some simple forecasting methods

Some forecasting methods are very simple and surprisingly effective.
Here are four methods that we will use as benchmarks for other
forecasting methods.

###Average method {-}

Here, the forecasts of all future values are equal to the mean of the
historical data. If we let the historical data be denoted by
$y_{1},\dots,y_{T}$, then we can write the forecasts as 
$$ {\hat{y}}_{T+h|T} = \bar{y} = (y_{1}+\dots+y_{T})/T. $$
The notation $\hat{y}_{T+h|T}$ is a short-hand for the estimate of $y_{T+h}$ based on the data $y_1,\dots,y_T$.

Although we have used time series notation here, this method can also be used for cross-sectional data (when we are predicting a value not included in the
data set). Then the prediction for values not observed is the average of
those values that have been observed. The remaining methods in this
section are only applicable to time series data.

```r
meanf(y, h) 
# y contains the time series
# h is the forecast horizon
```

###Naïve method {-}

This method is only appropriate for time series data. All forecasts are
simply set to be the value of the last observation. That is, the
forecasts of all future values are set to be $y_{T}$, where $y_T$
is the last observed value. This method works remarkably well for many
economic and financial time series.

```r
naive(y, h)
rwf(y, h) # Alternative
```

###Seasonal naïve method {-}

A similar method is useful for highly seasonal data. In this case, we
set each forecast to be equal to the last observed value from the same
season of the year (e.g., the same month of the previous year).
Formally, the forecast for time $T+h$ is written as 
$$ y_{T+h-km} \text{ where $m=$ seasonal period, $k=\lfloor (h-1)/m\rfloor+1$,} $$ 
and $\lfloor u \rfloor$ denotes the integer part of $u$. That looks more complicated than it really is. 
For example, with monthly data, the forecast for all future February
values is equal to the last observed February value. With quarterly
data, the forecast of all future Q2 values is equal to the last observed
Q2 value (where Q2 means the second quarter). Similar rules apply for
other months and quarters, and for other seasonal periods.

```r
snaive(y, h)
```


###Drift method {-}

A variation on the naïve method is to allow the forecasts to increase or
decrease over time, where the amount of change over time (called the
drift) is set to be the average change seen in the historical data. So
the forecast for time $T+h$ is given by 
$$ y_{T} + \frac{h}{T-1}\sum_{t=2}^T (y_{t}-y_{t-1}) = y_{T} + h \left( \frac{y_{T} -y_{1}}{T-1}\right). $$ 
This is equivalent to drawing a line between the first and last observation, and extrapolating it into
the future.

```r
rwf(y, h, drift=TRUE)
```

###Examples {-}

Figure \@ref(fig:beerf) shows the first three methods applied to the quarterly beer production data. 

```{r beerf, fig.cap="Forecasts of Australian quarterly beer production."}
beer2 <- window(ausbeer,start=1992,end=2006-.1)
beerfit1 <- meanf(beer2, h=11)
beerfit2 <- naive(beer2, h=11)
beerfit3 <- snaive(beer2, h=11)

plot(beerfit1, plot.conf=FALSE, 
  main="Forecasts for quarterly beer production")
lines(beerfit2$mean,col=2)
lines(beerfit3$mean,col=3)
legend("topright",lty=1,col=c(4,2,3),
  legend=c("Mean method","Naive method","Seasonal naive method"))
```

In Figure \@ref(fig:djf), the non-seasonal methods were applied to a series of 250 days of the Dow Jones Index. 

```{r djf, fig.cap="Forecasts based on 250 days of the Dow Jones Index."}
dj2 <- window(dj,end=250)
plot(dj2,main="Dow Jones Index (daily ending 15 Jul 94)",
  ylab="",xlab="Day",xlim=c(2,290))
lines(meanf(dj2,h=42)$mean,col=4)
lines(rwf(dj2,h=42)$mean,col=2)
lines(rwf(dj2,drift=TRUE,h=42)$mean,col=3)
legend("topleft",lty=1,col=c(4,2,3),
  legend=c("Mean method","Naive method","Drift method"))
```

Sometimes one of these simple methods will be the best forecasting
method available. But in many cases, these methods will serve as
benchmarks rather than the method of choice. That is, whatever
forecasting methods we develop, they will be compared to these simple
methods to ensure that the new method is better than these simple
alternatives. If not, the new method is not worth considering.

