## Evaluating forecast accuracy

###Forecast accuracy measures {-}

Let $y_{i}$ denote the $i$th observation and $\hat{y}_{i}$ denote a forecast of $y_{i}$.

### Scale-dependent errors {-}

The forecast error is simply $e_{i}=y_{i}-\hat{y}_{i}$, which is on the same scale as the data. Accuracy measures that are based on $e_{i}$ are therefore scale-dependent and cannot be used to make comparisons between series that are on different scales.

The two most commonly used scale-dependent measures are based on the absolute errors or squared errors:
\begin{align*}
  \text{Mean absolute error: MAE} & = \text{mean}(|e_{i}|),\\
  \text{Root mean squared error: RMSE} & = \sqrt{\text{mean}(e_{i}^2)}.
\end{align*}
When comparing forecast methods on a single data set, the MAE is popular as it is easy to understand and compute.

### Percentage errors {-}

The percentage error is given by $p_{i} = 100 e_{i}/y_{i}$. Percentage errors have the advantage of being scale-independent, and so are frequently used to compare forecast performance between different data sets. The most commonly used measure is:
$$
  \text{Mean absolute percentage error: MAPE} = \text{mean}(|p_{i}|).
$$
Measures based on percentage errors have the disadvantage of being infinite or undefined if $y_{i}=0$ for any $i$ in the period of interest, and having extreme values when any $y_{i}$ is close to zero. Another problem with percentage errors that is often overlooked is that they assume a meaningful zero. For example, a percentage error makes no sense when measuring the accuracy of temperature forecasts on the Fahrenheit or Celsius scales.

They also have the disadvantage that they put a heavier penalty on negative errors than on positive errors. This observation led to the use of the so-called "symmetric" MAPE (sMAPE) proposed by Armstrong (1985, p.348), which was used in the M3 forecasting competition. It is defined by
$$ 
  \text{sMAPE} = \text{mean}\left(200|y_{i} - \hat{y}_{i}|/(y_{i}+\hat{y}_{i})\right). 
$$
However, if $y_{i}$ is close to zero, $\hat{y}_{i}$ is also likely to be close to zero. Thus, the measure still involves division by a number close to zero, making the calculation unstable. Also, the value of sMAPE can be negative, so it is not really a measure of "absolute percentage errors" at all.

Hyndman and Koehler (2006) recommend that the sMAPE not be used. It is included here only because it is widely used, although we will not use it in this book.

### Scaled errors {-}

Scaled errors were proposed by Hyndman and Koehler (2006) as an alternative to using percentage errors when comparing forecast accuracy across series on different scales. They proposed scaling the errors based on the *training* MAE from a simple forecast method. For a non-seasonal time series, a useful way to define a scaled error uses naïve forecasts:
$$
  q_{j} = \frac{\displaystyle e_{j}}
    {\displaystyle\frac{1}{T-1}\sum_{t=2}^T |y_{t}-y_{t-1}|}.
$$

Because the numerator and denominator both involve values on the scale of the original data, $q_{j}$ is independent of the scale of the data. A scaled error is less than one if it arises from a better forecast than the average naïve forecast computed on the training data. Conversely, it is greater than one if the forecast is worse than the average naïve forecast computed on the training data. For seasonal time series, a scaled error can be defined using seasonal naïve forecasts: 
$$ 
  q_{j} = \frac{\displaystyle e_{j}}
    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}. 
$$

The *mean absolute scaled error* is simply
$$
  \text{MASE} = \text{mean}(|q_{j}|).
$$
Similarly, the *mean squared scaled error* (MSSE) can be defined where the errors (on the training data and test data) are squared instead of using absolute values.

###Examples {-}

```{r beeraccuracy, fig.cap="Forecasts of Australian quarterly beer production using data up to the end of 2007."}
beer2 <- window(ausbeer,start=1992,end=c(2007,4))
beerfit1 <- meanf(beer2,h=10)
beerfit2 <- rwf(beer2,h=10)
beerfit3 <- snaive(beer2,h=10)
tmp <- cbind(Data=window(ausbeer, start=1992),
             Mean=beerfit1$mean, 
             Naive=beerfit2$mean, 
             SeasonalNaive=beerfit3$mean)
autoplot(tmp) + xlab("Year") + ylab("Megalitres") +
  ggtitle("Forecasts for quarterly beer production") + 
  scale_color_manual(values=c('#000000','#1b9e77','#d95f02','#7570b3'),
                     breaks=c("Mean","Naive","SeasonalNaive"),
                     name="Forecast Method")
```


Figure \ref@(fig:beeraccuracy) shows three forecast methods applied to the quarterly Australian beer production using data only to the end of 2007. The actual values for the period 2008--2010 are also shown. We compute the forecast accuracy measures for this period.


```r
beer3 <- window(ausbeer, start=2008)
accuracy(beerfit1, beer3)
accuracy(beerfit2, beer3)
accuracy(beerfit3, beer3)
```

```{r echo=FALSE}
beer3 <- window(ausbeer, start=2008)
tab <- matrix(NA,ncol=4,nrow=3)
tab[1,] <- accuracy(beerfit1, beer3)[2,c(2,3,5,6)]
tab[2,] <- accuracy(beerfit2, beer3)[2,c(2,3,5,6)]
tab[3,] <- accuracy(beerfit3, beer3)[2,c(2,3,5,6)]
colnames(tab) <- c("RMSE","MAE","MAPE","MASE")
rownames(tab) <- c("Mean method", "Naïve method", "Seasonal naïve method")
knitr::kable(tab, digits=2)
```


It is obvious from the graph that the seasonal naïve method is best for these data, although it can still be improved, as we will discover later. Sometimes, different accuracy measures will lead to different results as to which forecast method is best. However, in this case, all the results point to the seasonal naïve method as the best of these three methods for this data set.

To take a non-seasonal example, consider the Dow Jones Index. The following graph shows the 250 observations ending on 15 July 1994, along with forecasts of the next 42 days obtained from three different methods.

```{r, fig.cap="Forecasts of the Dow Jones Index from 16 July 1994."}
dj2 <- window(dj, end=250)
djfc1 <- meanf(dj2, h=42)
djfc2 <- rwf(dj2, h=42)
djfc3 <- rwf(dj2, drift=TRUE, h=42)

tmp <- cbind(Data=dj, Mean=djfc1$mean, Naive=djfc2$mean, Drift=djfc3$mean)
autoplot(tmp) + xlab("Day") + ylab("") +
  ggtitle("Dow Jones Index (daily ending 15 Jul 94)") +
  scale_color_manual(values=c('#000000','#1b9e77','#d95f02','#7570b3'),
                     breaks=c("Mean","Naive","Drift"),
                     name="Forecast Method")
```

```r
dj3 <- window(dj, start=251)
accuracy(djfc1, dj3)
accuracy(djfc2, dj3)
accuracy(djfc3, dj3)
```

```{r echo=FALSE}
dj3 <- window(dj, start=251)
tab <- matrix(NA,ncol=4,nrow=3)
tab[1,] <- accuracy(djfc1, dj3)[2,c(2,3,5,6)]
tab[2,] <- accuracy(djfc2, dj3)[2,c(2,3,5,6)]
tab[3,] <- accuracy(djfc3, dj3)[2,c(2,3,5,6)]
colnames(tab) <- c("RMSE","MAE","MAPE","MASE")
rownames(tab) <- c("Mean method", "Naïve method", "Drift method")
knitr::kable(tab, digits=2)
```

Here, the best method is the drift method (regardless of which accuracy measure is used).

###Training and test sets {-}

It is important to evaluate forecast accuracy using genuine forecasts. That is, it is invalid to look at how well a model fits the historical data; the accuracy of forecasts can only be determined by considering how well a model performs on new data that were not used when fitting the model. When choosing models, it is common to use a portion of the available data for fitting, and use the rest of the data for testing the model, as was done in the above examples. Then the testing data can be used to measure how well the model is likely to forecast on new data.

The size of the test set is typically about 20% of the total sample, although this value depends on how long the sample is and how far ahead you want to forecast. The size of the test set should ideally be at least as large as the maximum forecast horizon required. The following points should be noted.

-   A model which fits the data well does not necessarily forecast well.
-   A perfect fit can always be obtained by using a model with enough parameters.
-   Over-fitting a model to data is as bad as failing to identify the systematic pattern in the data.

Some references describe the test set as the "hold-out set" because these data are "held out" of the data used for fitting. Other references call the training set the "in-sample data" and the test set the "out-of-sample data". We prefer to use "training set" and "test set" in this book.

###Cross-validation {-}

A more sophisticated version of training/test sets is cross-validation. For cross-sectional data, cross-validation works as follows.

1.  Select observation $i$ for the test set, and use the remaining observations in the training set. Compute the error on the test observation.
2.  Repeat the above step for $i=1,2,\dots,N$ where $N$ is the total number of observations.
3.  Compute the forecast accuracy measures based on the errors obtained.

This is a much more efficient use of the available data, as you only omit one observation at each step. However, it can be very time consuming to implement.

For time series data, the procedure is similar but the training set consists only of observations that occurred *prior* to the observation that forms the test set. Thus, no future observations can be used in constructing the forecast. However, it is not possible to get a reliable forecast based on a very small training set, so the earliest observations are not considered as test sets. Suppose $k$ observations are required to produce a reliable forecast. Then the process works as follows.

1.  Select the observation at time $k+i$ for the test set, and use the observations at times $1,2,\dots,k+i-1$ to estimate the forecasting model. Compute the error on the forecast for time $k+i$.
2.  Repeat the above step for $i=1,2,\dots,T-k$ where $T$ is the total number of observations.
3.  Compute the forecast accuracy measures based on the errors obtained.

This procedure is sometimes known as a "rolling forecasting origin" because the "origin" ($k+i-1$) at which the forecast is based rolls forward in time.

With time series forecasting, one-step forecasts may not be as relevant as multi-step forecasts. In this case, the cross-validation procedure based on a rolling forecasting origin can be modified to allow multi-step errors to be used. Suppose we are interested in models that produce good $h$-step-ahead forecasts.

1.  Select the observation at time $k+h+i-1$ for the test set, and use the observations at times $1,2,\dots,k+i-1$ to estimate the forecasting model. Compute the $h$-step error on the forecast for time $k+h+i-1$.
2.  Repeat the above step for $i=1,2,\dots,T-k-h+1$ where $T$ is the total number of observations.
3.  Compute the forecast accuracy measures based on the errors obtained.

When $h=1$, this gives the same procedure as outlined above.
