---
title: "White noise"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

## White noise

```{r, echo=TRUE, fig.height=2.7}
set.seed(3)
wn <- ts(rnorm(36))
autoplot(wn)
```

 * "White noise" is just a time series of iid data.

## White noise ACF

\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=2.7}
ggAcf(wn) + ggtitle("Sample ACF for white noise")
```

 * We expect each autocorrelation to be close to zero.
 * 95% of all autocorrelations for white noise should lie within the blue lines on the graph.
 * If this is not the case, the series is probably not white noise.


## Example: Pigs slaughtered

\fontsize{10}{11}\sf

```{r pigs, echo=FALSE}
pigs <- window(pigs, start=1990)
```

```{r, fig.height=3, echo=TRUE}
autoplot(pigs) +
  xlab("Year") + ylab("thousands") +
  ggtitle("Monthly number of pigs slaughtered in Victoria")
```

* Is this white noise?

## Example: Pigs slaughtered

\fontsize{10}{10}\sf

```{r, fig.height=2.5, echo=TRUE}
ggAcf(pigs) +
  ggtitle("ACF of monthly pigs slaughtered in Victoria")
```

 * Difficult to detect pattern in time plot.
 * ACF shows some significant autocorrelation at lags 1, 2, and 3.
 * These show the series is \alert{not a white noise series}.
 * There is some information in the data that can be used to forecast future values.

## Ljung-Box test

\begin{block}{}
The Ljung-Box test considers the first $h$ autocorrelation values together, rather than individually. A significant test (small $p$-value) indicates the data are probably not white noise.
\end{block}

\pause

\textbf{For the pigs data:}

```{r, ljungbox, echo=TRUE}
Box.test(pigs, lag=24, fitdf=0, type="Lj")
```

 * The small p-value suggests the data are probably not white noise.

## White noise

* White noise is a time series that is purely random.
* We can test for white noise by looking at an ACF plot or by doing a Ljung-Box test.
