---
title: "Welcome to Forecasting Using R"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


## What you will learn

  * Exploring and visualizing time series
  * Simple benchmark methods for forecasting
  * Exponential smoothing and ARIMA models
  * Advanced forecasting methods
  * Measuring forecast accuracy and choosing the best method

## Course textbook

**Hyndman, R. J. \& Athanasopoulos, G. (2017)**

**\emph{Forecasting: principles and practice}, 2nd edition**

\includegraphics[width=2.5cm]{fppcover.jpg}

  * Free and online at \alert{\textbf{\url{OTexts.org/fpp2/}}}
  * Data sets in associated R package `fpp2`
  * R code for all examples

## Time series data

\fontsize{10}{10}\sf

  - Daily IBM stock prices
  - Monthly rainfall in London
  - Quarterly Australian beer production
  - Annual Google profits

\pause


```{r, echo=FALSE, fig.height=2.8, cache=TRUE, warning=FALSE, message=FALSE}
auscafe %>% window(start=2000) %>% autoplot() +
  ggtitle("Monthly Australian expenditure on eating out") +
  xlab("Year") +
  ylab("A$ billion") +
  scale_x_continuous(breaks=seq(2000,2015,by=5),
                     minor_breaks=2001:2018)
```

\pause

\begin{alertblock}{}
\bfseries Forecasting is estimating how the sequence of\\ observations will continue into the future.
\end{alertblock}

## Forecasts of monthly Australian expenditure on eating out

```{r, echo=FALSE, fig.height=3, cache=TRUE, message=FALSE}
fit <- auto.arima(window(auscafe,start=2000), lambda=0, D=1)
fit %>% forecast(h=36) %>% autoplot(include=120) +
  ggtitle("3-year forecasts") +
  ylab("A$ billion") +
  xlab("Year") +
  scale_x_continuous(breaks=seq(2000,2025,by=5),
                     minor_breaks=2001:2020)
```

\pause

###
 * What forecasting methods are available that take account of trend, seasonality and other features of the data?
 * How to measure the accuracy of your forecasts?
 * How to choose a good forecasting model?

##

\centering\large\bfseries
Let's get started!
