---
title: "Trends, seasonality and cyclicity"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


## Time series patterns

\begin{tabular}{lp{9cm}}
\bf\alert{Trend}    & a pattern exists involving a long-term increase or decrease in the data.\\
\bf\alert{Seasonal} & a periodic pattern exists due to the calendar\\
                    & (e.g., the quarter, the month, or day of the week).\\
\bf\alert{Cyclic}   & a pattern exists where the data exhibits rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).
\end{tabular}


## Examples of time series patterns

```{r, message=FALSE}
autoplot(window(elec, start=1980)) +
  ggtitle("Australian electricity production") +
  xlab("Year") + ylab("GWh") +
  scale_x_continuous(breaks=seq(1980,1995,by=5),
                     minor_breaks=1980:1999)
```

## Examples of time series patterns

```{r}
autoplot(bricksq) +
  ggtitle("Australian clay brick production") +
  xlab("Year") + ylab("million units")
```

## Examples of time series patterns

```{r}
autoplot(ustreas) +
  ggtitle("US Treasury bill contracts") +
  xlab("Day") + ylab("price")
```

## Examples of time series patterns

```{r}
autoplot(lynx) +
  ggtitle("Annual Canadian lynx trappings") +
  xlab("Year") + ylab("Number trapped")
```

## Seasonal or cyclic?

**Differences between seasonal and cyclic patterns:**

 * seasonal pattern constant length; cyclic pattern variable length
 * average length of cycle longer than length of seasonal pattern
 * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

###
The timing of peaks and troughs is predictable with seasonal data,

but unpredictable in the long term with cyclic data.

