---
title: "Time series cross-validation"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


##Time series cross-validation {-}

**Traditional evaluation**

\begin{block}{}
\includegraphics[width=11cm]{figs/traintest-1.pdf}
\end{block}

\pause

**Time series cross-validation**

\begin{block}{}
\includegraphics[width=11cm]{figs/cv1-1.pdf}
\end{block}

\pause

 * Forecast accuracy averaged over test sets.
 * Also known as "forecast evaluation on a rolling origin"

##Time series cross-validation {-}

**Traditional evaluation**

\begin{block}{}
\includegraphics[width=11cm]{figs/traintest-1.pdf}
\end{block}

**Time series cross-validation**

\begin{block}{}
\includegraphics[width=11cm]{figs/cv2-1.pdf}
\end{block}

 * Forecast accuracy averaged over test sets.
 * Also known as "forecast evaluation on a rolling origin"


##Time series cross-validation {-}

**Traditional evaluation**

\begin{block}{}
\includegraphics[width=11cm]{figs/traintest-1.pdf}
\end{block}

**Time series cross-validation**

\begin{block}{}
\includegraphics[width=11cm]{figs/cv3-1.pdf}
\end{block}

 * Forecast accuracy averaged over test sets.
 * Also known as "forecast evaluation on a rolling origin"


## tsCV function

\small

**RMSE using time series cross-validation**
```{r tscv, cache=TRUE, echo=TRUE}
e <- tsCV(oil, forecastfunction=naive, h=1)
sqrt(mean(e^2, na.rm=TRUE))
```

 * When there are no parameters to be estimated, `tsCV` with `h=1` will give the same values as `residuals`.


## tsCV function

\fontsize{10}{10}\sf

```{r tscv2, echo=TRUE}
sq <- function(u){u^2}
for(h in 1:10)
{
  oil %>% tsCV(forecastfunction=naive, h=h) %>%
    sq() %>% mean(na.rm=TRUE) %>% print()
}
```

* The RMSE increases with the forecast horizon.

## Time series cross-validation

\large

* Choose the model with the smallest RMSE computed using time series cross-validation.
* Compute it at the forecast horizon of most interest to you.
