---
title: "Forecasts and potential futures"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

## Sample futures

```{r austa, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
fit <- ets(austa)
df <- cbind(austa, simulate(fit,10))
for(i in seq(99))
  df <- cbind(df, simulate(fit,10))
colnames(df) <- c("Data", paste("Future",1:100))
set.seed(1)
cols <- scale_colour_manual(values=c('#000000',rainbow(5)[-2],sample(rainbow(95))),
                     breaks=c("Data",paste("Future",1:100)),
                     name=" ")
```

```{r austa0, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(austa) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  ylim(.85,10.0) + xlim(1981,2025)
```

## Sample futures

```{r austa1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:2]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```


## Sample futures

```{r austa2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:3]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Sample futures

```{r austa3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:4]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Sample futures

```{r austa4, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:5]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Sample futures

```{r austa5, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:6]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Sample futures

```{r austa15, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:15]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Sample futures

```{r austa100, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(df[,1:100]) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
  cols + ylim(.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```


## Forecast intervals

```{r austapi, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(forecast(fit)) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Forecasts of total international visitors to Australia") +
  ylim(0.85,10.0) + xlim(1981,2025) +
  theme(legend.position="none")
```

## Forecast intervals

\fontsize{10}{11}\sf

```{r austapi2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
autoplot(forecast(fit)) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Forecasts of total international visitors to Australia") +
  ylim(0.85,10.0) + xlim(1981,2025)
```

* The 80% forecast intervals should contain 80% of the future observations.
* The 95% forecast intervals should contain 95% of the future observations.

##
\centering\Large\bfseries
Time to start forecasting!
