---
title: "Fitted values and residuals"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

## Fitted values and residuals

A \alert{fitted value} is the forecast of an observation using all previous observations.

 - That is, they are one-step forecasts.
 - Often not true forecasts since parameters are estimated on all data.

\pause

A \alert{residual} is the difference between an observation and its fitted value.

 - That is, they are one-step forecast errors.

## Example: oil production

```{r oil, fig.height=3, warning=FALSE, echo=TRUE}
fc <- naive(oil)
autoplot(oil, series="Data") + xlab("Year") +
  autolayer(fitted(fc), series="Fitted") +
  ggtitle("Oil production in Saudi Arabia")
```

## Example: oil production

```{r dj2, echo=TRUE}
autoplot(residuals(fc))
```

## Residuals should look like white noise

  1. They should be uncorrelated.
  2. They should have mean zero.

\pause

**Useful properties (for computing prediction intervals)**

  3. They should have constant variance.
  4. They should be normally distributed.

\pause


\begin{alertblock}{}
We can test these assumptions using the \texttt{checkresiduals()} function.
\end{alertblock}

## Example: oil production

\fontsize{9}{9}\sf

```{r, warning=FALSE, echo=TRUE, fig.height=2.5}
checkresiduals(fc)
```
