---
title: "Exponential smoothing methods with trend"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##Holt's linear trend

\begin{block}{Simple exponential smoothing}\vspace*{-.6cm}
\begin{align*}
\text{Forecast }&& \hat{y}_{t+h|t} &= \ell_{t} \\
\text{Level }&& \ell_{t} &= \alpha y_{t} + (1 - \alpha)\ell_{t-1} \hspace*{1.8cm}
\end{align*}
\end{block}\pause

\begin{block}{Holt's linear trend}\vspace*{-.6cm}
\begin{align*}
\text{Forecast }&& \hat{y}_{t+h|t} &= \ell_{t} + hb_{t} \\
\text{Level }&& \ell_{t} &= \alpha y_{t} + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
\text{Trend }&& b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 -\beta^*)b_{t-1},
\end{align*}
\end{block}\pause

  * Two smoothing parameters $\alpha$ and $\beta^*$ ($0\le\alpha,\beta^*\le1$).
  * Choose $\alpha, \beta^*, \ell_0, b_0$ to minimise SSE.

##Holt's method in R

```{r}
airpassengers <- window(ausair, start=1990, end=2004)
```


```{r, fig.height=3., echo=TRUE}
airpassengers %>% holt(h=5) %>% autoplot()
```

##Damped trend method
\begin{block}{Component form}\vspace*{-0.6cm}
\begin{align*}
\hat{y}_{t+h|t} &= \ell_{t} + (\phi+\phi^2 + \dots + \phi^{h})b_{t} \\
\ell_{t} &= \alpha y_{t} + (1 - \alpha)(\ell_{t-1} + \phi b_{t-1})\\
b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 -\beta^*)\phi b_{t-1}.
\end{align*}
\end{block}
\pause

  * Damping parameter $0<\phi<1$.
  * If $\phi=1$, identical to Holt's linear trend.
  * Short-run forecasts trended, long-run forecasts constant.

##Example: Air passengers

\fontsize{10}{11}\sf

```{r, echo=TRUE, fig.height=3}
fc1 <- holt(airpassengers, h=15, PI=FALSE)
fc2 <- holt(airpassengers, damped=TRUE, h=15, PI=FALSE)
autoplot(airpassengers) + xlab("Year") + ylab("millions") +
  autolayer(fc1, series="Linear trend") +
  autolayer(fc2, series="Damped trend")
```

