---
title: "Exponential smoothing methods with trend and seasonality"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


##Holt-Winters additive method

\begin{block}{Holt-Winters additive method}\vspace*{-0.6cm}
\begin{align*}
\hat{y}_{t+h|t} &= \ell_{t} + hb _{t} + s_{t-m+h_{m}^{+}} \\
\ell_{t} &= \alpha(y_{t} - s_{t-m}) + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 - \beta^*)b_{t-1}\\
s_{t} &= \gamma (y_{t}-\ell_{t-1}-b_{t-1}) + (1-\gamma)s_{t-m}
\end{align*}
\end{block}\vspace*{3.35ex}


  * $s_{t-m+h_{m}^{+}}=$ seasonal component from final year of data.
  * Smoothing parameters:  $0\le \alpha\le 1$, \sspace $0\le \beta^*\le 1$, \sspace $0\le \gamma\le 1-\alpha$.
  * $m=$  period of seasonality (e.g. $m=4$ for quarterly data).
  * seasonal component averages zero

\vspace*{10cm}

##Holt-Winters multiplicative method

\begin{block}{Holt-Winters multiplicative method}\vspace*{-0.6cm}
    \begin{align*}
        \hat{y}_{t+h|t} &= (\ell_{t} + hb_{t})s_{t-m+h_{m}^{+}} \\
        \ell_{t} &= \alpha \frac{y_{t}}{s_{t-m}} + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
        b_{t} &= \beta^*(\ell_{t}-\ell_{t-1}) + (1 - \beta^*)b_{t-1}        \\
        s_{t} &= \gamma \frac{y_{t}}{(\ell_{t-1} + b_{t-1})} + (1 - \gamma)s_{t-m}
    \end{align*}
\end{block}

  * $s_{t-m+h_{m}^{+}}=$ seasonal component from final year of data.
  * Smoothing parameters:  $0\le \alpha\le 1$, \sspace $0\le \beta^*\le 1$, \sspace  $0\le \gamma\le 1-\alpha$.
  * $m=$  period of seasonality (e.g. $m=4$ for quarterly data).
  * seasonal component averages one.

\vspace*{10cm}

##Example: Visitor Nights

```{r 7-HW, echo=TRUE}
aust <- window(austourists, start=2005)
fc1 <- hw(aust, seasonal="additive")
fc2 <- hw(aust, seasonal="multiplicative")
```

```{r, fig.height=3.2}
tmp <- cbind(Data=aust,
  "HW additive forecasts" = fc1[["mean"]],
  "HW multiplicative forecasts" = fc2[["mean"]])

autoplot(tmp) + xlab("Year") +
  ylab("millions") + ggtitle("International visitor night in Australia") +
  scale_color_manual(name="",
    values=c('#000000','#1b9e77','#d95f02'),
    breaks=c("Data","HW additive forecasts","HW multiplicative forecasts"))
```


## Taxonomy of exponential smoothing methods

\fontsize{12}{13}\sf

\begin{block}{}
\begin{tabular}{ll|ccc}
& &\multicolumn{3}{c}{\bf Seasonal Component} \\
\multicolumn{2}{c|}{\bf Trend}& N & A & M\\
\multicolumn{2}{c|}{\bf Component}  & (None)    & (Additive)  & (Multiplicative)\\
\cline{3-5} &&&&\\[-0.4cm]
N & (None) & (N,N) & (N,A) & (N,M)\\
&&&&\\[-0.4cm]
A & (Additive) & (A,N) & (A,A) & (A,M)\\
&&&&\\[-0.4cm]
A\damped & (Additive damped) & (A\damped,N) & (A\damped,A) & (A\damped,M)
\end{tabular}
\end{block}\fontsize{12}{14}\sf

\begin{tabular}{lp{7.5cm}l}
\alert{(N,N)}:        & Simple exponential smoothing               & \texttt{ses()}\\
\alert{(A,N)}:        & Holt's linear method                       & \texttt{holt()}\\
\alert{(A\damped,N)}: & Additive damped trend method               & \texttt{hw()}\\
\alert{(A,A)}:~~      & Additive Holt-Winters' method              & \texttt{hw()}\\
\alert{(A,M)}:        & Multiplicative Holt-Winters' method        & \texttt{hw()}\\
\alert{(A\damped,M)}: & Damped multiplicative Holt-Winters' method & \texttt{hw()}
\end{tabular}

\begin{block}{}\fontsize{12}{14}\sf
There are also multiplicative trend methods (not recommended).
\end{block}


