---
title: "State space models for exponential smoothing"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##ETS models

   * Each exponential smoothing method can be written as an "innovations state space model" in two ways.
   * ETS(Error, Trend, Seasonal):
      * Error $=\{$A,M$\}$
      * Trend $=\{$N,A,A\damped$\}$
      * Seasonal $=\{$N,A,M$\}$.
  * All parameters can be estimated using the "likelihood":  the probability of the data arising from the specified model.
  * For models with additive errors, this is equivalent to minimising SSE.
  * Choose the best model by minimizing the AICc (asymptotically equivalent to time series cross-validation).


##Example: Australian air traffic
\fontsize{10}{10}\sf

```{r, echo=TRUE, fig.height=4}
ets(ausair)
```

##Example: Australian air traffic

```{r, echo=TRUE, fig.height=4}
ausair %>% ets() %>% forecast() %>% autoplot()
```

##Example: Monthly cortecosteroid drug sales
\fontsize{8}{8}\sf

```{r, echo=TRUE}
ets(h02)
```

##Example: Monthly cortecosteroid drug sales

```{r, echo=TRUE, fig.height=4}
h02 %>% ets() %>% forecast() %>% autoplot()
```
