---
title: "Transformations for variance stabilization"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


##Variance stabilization

If the data show increasing variation as the level of the series increases, then a transformation can be useful.

Denote original observations as $y_1,\dots,y_n$ and transformed
observations as $w_1, \dots, w_n$.

\begin{block}{\footnotesize Mathematical transformations for stabilizing
variation}
\begin{tabular}{llc}
Square root & $w_t = \sqrt{y_t}$ & $\downarrow$ \\[0.2cm]
Cube root & $w_t = \sqrt[3]{y_t}$ & Increasing \\[0.2cm]
Logarithm & $w_t = \log(y_t)$  & strength \\[0.2cm]
Inverse   & $w_t = -1/y_t$& $\downarrow$
\end{tabular}
\end{block}
\pause

Logarithms are useful because they are more interpretable:
changes in a log value are **relative (percent) changes on the original
scale.**

##Variance stabilization

```{r, echo=TRUE}
autoplot(usmelec) +
  xlab("Year") + ylab("") +
  ggtitle("US monthly net electricity generation")
```

##Variance stabilization

```{r, echo=TRUE}
autoplot(usmelec^0.5) +
  xlab("Year") + ylab("") +
  ggtitle("Square root electricity generation")
```

##Variance stabilization

```{r, echo=TRUE}
autoplot(usmelec^0.33333) +
  xlab("Year") + ylab("") +
  ggtitle("Cube root electricity generation")
```

##Variance stabilization

```{r, echo=TRUE}
autoplot(log(usmelec)) +
  xlab("Year") + ylab("") +
  ggtitle("Log electricity generation")
```

##Variance stabilization

```{r, echo=TRUE}
autoplot(-1/usmelec) +
  xlab("Year") + ylab("") +
  ggtitle("Inverse electricity generation")
```

##Box-Cox transformations

Each of these transformations is close to a member of the
family of \textbf{Box-Cox transformations}:
$$w_t = \left\{\begin{array}{ll}
        \log(y_t),      & \quad \lambda = 0; \\
        (y_t^\lambda-1)/\lambda ,         & \quad \lambda \ne 0.
\end{array}\right.
$$\pause

* $\lambda=1$: No substantive transformation
* $\lambda=\frac12$: Square root plus linear transformation
* $\lambda=\frac13$: Cube root plus linear transformation
* $\lambda=0$: Natural logarithm
* $\lambda=-1$: Inverse transformation

\pause


###Automated Box-Cox transformations

```{r, echo=TRUE}
(BoxCox.lambda(usmelec))
```


##Back-transformation

\fontsize{10}{13}\sf

```{r,echo=TRUE}
usmelec %>%
  ets(lambda=-0.57) %>%
  forecast(h=60) %>%
  autoplot()
```

