---
title: "ARIMA models"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##ARIMA models

###Autoregressive (AR) models:
\centerline{$
  y_{t}  =  c  +  \phi_{1}y_{t - 1}  +  \phi_{2}y_{t - 2}  +  \cdots  +  \phi_{p}y_{t - p}  + e_{t},\qquad e_t \sim \text{white noise}.
$}
Multiple regression with \textbf{lagged observations} as predictors.

\pause

###Moving Average (MA) models:
\centerline{$
  y_{t}  =  c +  e_t + \theta_{1}e_{t - 1}  +  \theta_{2}e_{t - 2}  +  \cdots  + \theta_{q}e_{t - q},\qquad e_t \sim \text{white noise}.
$}
Multiple regression with \textbf{lagged \emph{errors}} as predictors.

\pause

###Autoregressive Moving Average (ARMA) models:
\centerline{$
y_{t}  =  c  +  \phi_{1}y_{t - 1}  +  \cdots  +  \phi_{p}y_{t - p}  + \theta_{1}e_{t - 1} +  \cdots  + \theta_{q}e_{t - q} +  e_{t}.
$}
Multiple regression with both \textbf{lagged observations} and \textbf{lagged errors} as predictors.

\pause

###ARIMA($p,d,q$) model:
* Combine ARMA model with $d$ lots of **differencing**.


##US net electricity generation

```{r}
autoplot(usnetelec) +
  xlab("Year") +
  ylab("billion kwh") +
  ggtitle("US net electricity generation")
```

##US net electricity generation

\fontsize{10}{11}\sf

```{r, echo=FALSE}
options(digits=4)
```

```{r, echo=TRUE}
fit <- auto.arima(usnetelec)
summary(fit)
```

##US net electricity generation

```{r, echo=TRUE, fig.height=4}
fit %>% forecast() %>% autoplot()
```


##How does auto.arima() work?

###Hyndman-Khandakar algorithm:

  * Select no.\ differences \textcolor{Orange}{$d$} via *unit root tests*.
  * Select \textcolor{Orange}{$p$} and \textcolor{Orange}{$q$} by minimising *AICc*.
  * Estimate parameters using *maximum likelihood estimation*.
  * Use stepwise search to traverse model space, to save time.

###Prediction intervals

* Prediction intervals *increase in size with forecast horizon*.
* Calculations assume residuals are *uncorrelated* and *normally distributed*.


