---
title: "Seasonal ARIMA models"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```


##Seasonal ARIMA models

\begin{block}{}\centering
\begin{tabular}{rcc}
ARIMA & \(~\underbrace{(p, d, q)}\) &
\(\underbrace{(P, D, Q)_{m}}\)\tabularnewline
& \({\uparrow}\) & \({\uparrow}\)\tabularnewline
& Non-seasonal part & Seasonal part of\tabularnewline
& of the model & of the model\tabularnewline
\end{tabular}
\end{block}

 * $p=$ number of ordinary AR lags ($y_{t-1},y_{t-2},\dots,y_{t-p}$)
 * $d=$ number of lag-1 differences
 * $q=$ number of ordinary MA lags ($\varepsilon_{t-1},\varepsilon_{t-2},\dots,\varepsilon_{t-q}$)
 * $P=$ number of seasonal AR lags ($y_{t-m},y_{t-2m},\dots,y_{t-Pm}$)
 * $D=$ number of seasonal differences
 * $Q=$ number of seasonal AR lags ($\varepsilon_{t-m},\varepsilon_{t-2m}\dots,\varepsilon_{t-Qm}$)
 * $m=$ number of observations per year.

\pause

###
A seasonal ARIMA model is non-linear

## Example: Monthly retail debit card usage in Iceland

```r
autoplot(debitcards) +
  xlab("Year") + ylab("million ISK") +
  ggtitle("Retail debit card usage in Iceland")
```

```{r, cache=TRUE, echo=FALSE, message=FALSE}
autoplot(debitcards) +
  xlab("Year") + ylab("million ISK") +
  ggtitle("Retail debit card usage in Iceland") +
  scale_x_continuous(breaks=seq(2000,2015,by=5),
    minor_breaks=2000:2020)
```

## Example: Monthly retail debit card usage in Iceland

```{r, cache=TRUE, echo=TRUE}
fit <- auto.arima(debitcards, lambda = 0)
fit
```

## Example: Monthly retail debit card usage in Iceland

```{r, echo=TRUE}
fit %>% forecast(h=36) %>% autoplot() + xlab("Year")
```

