---
title: "Dynamic harmonic regression"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##Dynamic harmonic regression

Periodic seasonality can be handled using pairs of Fourier terms:
$$
s_{k}(t) = \sin\left(\frac{2\pi k t}{m}\right)\qquad c_{k}(t) = \cos\left(\frac{2\pi k t}{m}\right)
$$

\begin{block}{}
$$
y_t = \beta_0 + \sum_{k=1}^K \left[\alpha_k s_k(t) + \gamma_k c_k(t)\right] + e_t$$
\end{block}

  * $m=$ seasonal period.
  * Every periodic function can be approximated by sums of sin and cos terms for large enough $K$.
  * Regression coefficients: $\alpha_k$ and $\gamma_k$.
  * $e_t$ can be modelled as a non-seasonal ARIMA process.
  * Assumes seasonal pattern is unchanging.

##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```{r, echo=FALSE}
cafe <- window(auscafe, start=2005)
```

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=1),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=1, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=1),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=1, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",1,"   AICC=",round(fit$aicc,2)))
```

##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=2),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=2, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=2),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=2, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",2,"   AICC=",round(fit$aicc,2)))
```

##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=3),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=3, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=3),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=3, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",3,"   AICC=",round(fit$aicc,2)))
```

##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=4),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=4, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=4),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=4, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",4,"   AICC=",round(fit$aicc,2)))
```


##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=5),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=5, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=5),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=5, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",5,"   AICC=",round(fit$aicc,2)))
```


##Example: Australian cafe expenditure
\fontsize{10}{13}\sf

```r
fit <- auto.arima(cafe, xreg=fourier(cafe, K=6),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=6, h=24)) %>%
  autoplot() + ylim(1.6,5.1)
```

```{r, echo=FALSE}
fit <- auto.arima(cafe, xreg=fourier(cafe, K=6),
                  seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=6, h=24)) %>%
  autoplot() + ylim(1.6,5.1) +
  xlab(paste("K=",6,"   AICC=",round(fit$aicc,2)))
```


##Dynamic harmonic regression


\begin{block}{}
$$
y_t = \beta_0 + \beta_1x_{t,1} + \dots + \beta_{t,r} x_{t,r}+ \sum_{k=1}^K \left[\alpha_k s_k(t) + \gamma_k c_k(t)\right] + e_t$$
\end{block}


* Other predictor variables can be added as well ($x_{t,1},\dots,x_{t,r}$)
* This is particularly useful for weekly data, daily data and sub-daily data.
* Choose $K$ to minimize the AICc.
* $K$ can not be more than $m/2$.

