---
title: "TBATS models"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##TBATS models

\begin{block}{\Large TBATS}
\textbf{\Large T}rigonometric terms for seasonality\\
\textbf{\Large B}ox-Cox transformations for heterogeneity\\
\textbf{\Large A}RMA errors for short-term dynamics\\
\textbf{\Large T}rend (possibly damped)\\
\textbf{\Large S}easonal (including multiple and non-integer periods)
\end{block}

## US Gasoline data

```{r, cache=TRUE, echo=TRUE}
gasoline %>% tbats() %>% forecast() %>%
  autoplot() +
  xlab("Year") + ylab("thousand barrels per day")
```

## Call centre data

```{r, cache=TRUE, echo=TRUE}
calls %>% window(start=20) %>%
  tbats() %>% forecast() %>%
  autoplot() + xlab("Weeks") + ylab("Calls")
```

##TBATS model

\begin{block}{\Large TBATS}\fontsize{13.5}{19}\sf
\textbf{\Large T}rigonometric terms for seasonality\\
\textbf{\Large B}ox-Cox transformations for heterogeneity\\
\textbf{\Large A}RMA errors for short-term dynamics\\
\textbf{\Large T}rend (possibly damped)\\
\textbf{\Large S}easonal (including multiple and non-integer periods)
\end{block}\vspace*{0.1cm}\fontsize{13}{15}\sf

* Handles non-integer seasonality, multiple seasonal periods.
* Entirely automated
* Prediction intervals often too wide
* Very slow on long series

