---
title: "Dynamic regression"
author: "Rob J Hyndman"
output:
  beamer_presentation:
    fig_height: 3
    fig_width: 6
    highlight: tango
    keep_tex: no
    theme: "metropolis"
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fpp2)
```

##Dynamic regression

\begin{block}{Regression model with ARIMA errors}\vspace*{-0.2cm}
\[
y_t = \beta_0 + \beta_1 x_{1,t} + \dots + \beta_r x_{r,t} + e_t,
\]\end{block}

  * $y_t$ modeled as function of $r$ explanatory variables
$x_{1,t},\dots,x_{r,t}$.
  * In ordinary regression, we assume that $e_t$ is white noise.
  * In dynamic regression, we allow $e_t$ to be an ARIMA process.


##US personal consumption and income

\fontsize{9}{13}\sf

```{r usconsump, echo=TRUE}
autoplot(uschange[,1:2], facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("Quarterly changes in US consumption and personal income")
```

##US personal consumption and income

\fontsize{9}{13}\sf

```{r, echo=TRUE}
ggplot(aes(x=Income,y=Consumption), data=as.data.frame(uschange)) +
  geom_point() +
  ggtitle("Quarterly changes in US consumption and personal income")
```


##Dynamic regression model for US personal consumption
\fontsize{10}{14}\sf

```{r usconsump2, echo=TRUE, fig.height=3}
fit <- auto.arima(uschange[,"Consumption"],
          xreg=uschange[,"Income"])
fit
```

##Residuals from dynamic regression model
\fontsize{9}{9}\sf

```{r , echo=TRUE, fig.height=2.5}
checkresiduals(fit)
```

##Forecasts from dynamic regression model
\fontsize{11}{13}\sf

```{r usconsump3, echo=TRUE, fig.height=3.}
fcast <- forecast(fit, xreg = rep(0.8, 8))
autoplot(fcast) +
  xlab("Year") + ylab("Percentage change")
```
